---
title: "Homework 3"
author: "Nidhi Patel"
date: "10/6/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6,
  out.width = "90%")

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

#### Problem 1

###### Describe data

```{r dataimport}
data("instacart")
```

Instacart dataset logs `r nrow(instacart)` observations over `r ncol(instacart)` Instacart online grocery order variables.  It not only keeps track of the user, order products, reorders, day and times of orders, but also organizes each product into departments and aisles they would be in store. For example, the product, Grated Pecorino Romano Cheese, is found in the specialty cheese aisle, which falls into the dairy eggs department.   The least ordered department is bulk, with 	1,359 orders, while the most ordered department is produce with 409,087 orders.

###### How many aisles? What is the most ordered aisle?

```{r P1count}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

There are 134 aisles. The fresh vegetables aisle is most ordered from with 150609 orders.

###### Plot aisle orders

```{r P1plot}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate( #this reorders the x axis.
    aisle = factor(aisle), #make it a factor
    aisle = fct_reorder(aisle, n) 
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  labs(
    y = "number of items",
    title = "number of items ordered in each aisle"
) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

This plot resembles an exponential graph because the most ordered from aisles are much greater than the others.  I find it interesting (and nice) that the most ordered items are fresh vegetables, followed by fresh fruit. Coming in third are packaged fruits and vegetables! Instacart users seem to like their vegetables!!

###### Top orders from baking ingredients, dog food care and packaged vegetable fruits aisles
 
```{r P1table1}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% #so, the order of aisle, product_name matters?
  summarize(n = n()) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

Light brown sugar, Snack Sticks Chicken & Rice Recipe dog treats and organic baby spinach are the most ordered items from baking ingredients, dog food care and packaged vegetable fruits aisles. Packaged fruits and vegetables are more ordere than the other two.  Dog food care is the least ordered from aise; maybe dog lovers probably buy treats while walking their dogs!!

###### When are apples and ice cream typically ordered?

```{r P1table2}
dow = tibble(
    order_dow = 0:6,
    day = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thurday", "Friday", "Saturday")
  )

 instacart = left_join(
    instacart, 
    dow,
    by = "order_dow"
  )

instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, day, order_dow) %>% 
  summarize(mean_hod = mean(order_hour_of_day)) %>% 
  arrange(order_dow) %>% 
  select(product_name, day, mean_hod) %>% 
  pivot_wider(
    names_from = "day",
    values_from = "mean_hod"
  ) %>% 
  knitr::kable()
```

Instacart users tend to order pink lady apples earlier in the day than ice cream. 

#### Problem 2

###### Tidy and data wrangle

```{r P2tidy}
accel =
  read_csv("./accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity"
  ) %>% 
  mutate(
    day = factor(day),
    minute = as.integer(minute),
    week = factor(week),
    day_id = factor(day_id),
    weekday = ifelse(day == "Saturday" | day == "Sunday", rep(FALSE), TRUE)
    )
```

The original version of this dataset contained 1436 columns and 25 rows. Activity was spread across 1440 columns and subsequently tidied into columns: minutes and activities. The resulting data has `r nrow(accel)` rows with `r ncol(accel)` activity observations. The variables 'week', 'day_id', and 'day' were converted to factors; 'minute' was converted to an integer. These were converted to factors to ease further wrangling and visualization. Activity was kept as an integer. Finally a logical variable of 'weekday' was added to differentiate weekdays from weekends. 

###### Total activity variable for each day 

```{r P2table}
activity =
  accel %>% 
  group_by(day, week) %>% 
  summarize(tot_act = sum(activity)) %>% 
  pivot_wider(
    names_from = "day",
    values_from = "tot_act"
  ) %>% 
  select(week, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday, Sunday)
knitr::kable(activity) #it wouldn't display the table when i knit if i piped and left () blank.
```

A pattern may be more consistent activity during the weekdays opposed to weekend, which varied greatly week by week.  I think it's interesting that the last two Saturday's recorded the same exact amount of activity and they it was extremely lower than anything else recorded. I wonder if this outlier is a mistake?

###### Plot 24-hour activity time

```{r P2plot}
accel %>% 
  group_by(day) %>% 
  ggplot(aes(x = minute, y = activity, color = week)) +
  geom_line(alpha = 0.1, size = 0.5) +
  stat_smooth(alpha = 0.2, size = 0.7) +
  scale_x_continuous(
    breaks = c(360, 720, 1080, 1440),
    labels = c("6h", "12h", "18h", "24h")
  ) +
  labs(
    title = "24 hour activity for each day",
    x = "hour of day"
  )
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

There are 35 x 1440 points of data to this plot, so it is difficult to differentiate day by day. However, by the spikes and lulls of activity, it looks like there is less activity in the morning and more midday and evening. Maybe this can be attributed to sleeping during the early morning?


#### Problem 3

###### Describe data

```{r P3describe}
library(p8105.datasets)
data("ny_noaa")
```

This dataset covers nearly 30 years of weather in New York state with `r ncol(ny_noaa)` columns of weather station, date, precipitation, snowfall, snow depth, maximum and minimum temperature. There are a total of `r nrow(ny_noaa)` observations.  There are quite a new missing data points, especially in minimum and maximum temperature.  It seems that this issue is pertinent to certain weather stations.

###### Data cleaning

```{r P3snowobs}
ny_noaa = ny_noaa %>%
separate(date, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(
    tmin = (as.integer(tmin) / 10),
    tmax = (as.integer(tmax) / 10),
    prcp = prcp / 10)

mostsnow = ny_noaa %>% 
  count(snow, sort = TRUE)
```

The most commonly observed value is 0 mm of snow. This makes a lot of sense because it only snows a quarter of the year! The second and third most recorded values are 25 mm and 13 mm, respectively. 

###### Average max temp in jan and july across years

```{r P3maxtemp}
 tmax_noaa = 
  ny_noaa %>% 
  filter(month %in% c("1", "7")) %>% 
  mutate(
    month = recode(
      month,
      `1` = "January",
      `7` = "July")) %>% 
  group_by(month, year, id) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = year, y = mean_tmax, color = id, alpha = 0.3)) + 
  geom_path() +
  labs(
    title = "average ax temperature in January in July across various NYS weather stations",
    y = "max temperature"
  ) +
  facet_grid(month ~ .) +
  theme(legend.position = "none")
```

There are many many weather stations across New York in this dataset, and therefore it is difficult to see each individual station.  However, the intense overlapping does allow us to see the range in max temperatures across the state, which looks pretty consistent through the years. The range in temperature differs more in January across the state than in July. In January, there is an outlier in 1982 and  a few in the 2000's. The most distinct outlier in the plot in July 1987 : the temp dropped below 15C!! 

###### tmax vs tmin

```{r P3patch}
library(patchwork)
library(ggridges)

tvt = ny_noaa %>% 
  drop_na(tmin, tmax) %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex() +
  theme(legend.position = "right") +
  labs(
    title = "min vs max temperatures",
    x = "min temp (C)",
    y = "max temp (C)"
  )

distrib = ny_noaa %>% 
  filter(snow > 0 & snow < 100) %>% 
  ggplot(aes(x = snow, y = year, group = year)) + 
  geom_density_ridges(scale = 0.85) +
  labs(
    title = "distribution of snowfall",
    x = "snow(mm)"
  )

tvt + distrib
```

"Min vs max temperatures" shows the most common minimum and maximum temperature differences in New York. The most frequently recurring temperatures in a day are around ~23C and ~7C.  I find most interesting min/max combination to be 60C and -29C, what a range!! The "distribution of snowfall" graph is also interesting as the last 5 years (2005-2010), have had much fewer snowfalls of 70-80mm than the years prior.  It is also interesting to see that there is a dip in snowfall at 20mm every year!! 
